---
title: "Constructing PGLS dataset for CCA species with functional and life-history trait data"
author: "Anne-Cecile Colin"
date: "10/06/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

librarian::shelf(tidyverse, readr, conflicted, ggtree, ape, data.table, DesiQuintans/desiderata)

#conflict_prefer("filter", "dplyr")
```

## 1. Aim

This project aims to construct the dataset for Victoria's PGLS analysis on all CCA species.
Victoria works on traits associated with drought in all species found at CCA which includes Angophora, Corymbia and Eucalyptus species.
Victoria's analysis aims to see whether P50 (and other functional/life-history traits) is different between humid/arid biomes, accounting for  the phylogeny.
To control for the phylogeny, i.e. if evolutionary relationships influence traits value, I am using the most resloved and complete phylogeny for this work produced by Thornhill et al., 2016.

In this project, I compare the species list from the Thornhill phylogeny, the CCA functional traits dataframe and the life-history traits extracted from EUCLID.

## 2. Read input data

``` {r reading raw data}

CCAnumber <- 
  # reading the dataset with the arboretum field ID (CCA#) and mother tree reference (UniqueID)
  read.csv("../data/CCA_ID.csv", header = T) %>%
  glimpse()

#colnames(CCAnumber)[1] <- "name"
nrow(CCAnumber)

species <- 
  # reading Victoria's species dataset (which contains all species from CCA)
  read.csv("../data/CCA_VPM_species_PGLS.csv", header = T) #%>%
  # keeping only one row per species-provenance (to avoid duplicates later on)
  # as it is a species-scale analysis, but Victoria will use =/= provenances per species
  distinct(UniqueID, .keep_all = TRUE)

species <- 
  #removing varieties and hybrids from the list
  species[!(species$Taxonresolution_ALA=="Hybrid") | species$Taxonresolution_ALA=="Variety",]
colnames(species)[4] <- "name"

view(species)
nrow(species)
```

## 3. Merging CCA datasets

``` {r merging dataframes}
species.ds <- 
  # merging the species dataset to CCA# by RowTree number (unique ID compared to merging by species name)
  # keeps all provenances per species and avoid mismatches by differences in species name column
  # such as symbols, lower case/caps, etc.
  left_join(species, CCAnumber, by = 'RowTree') #%>%
  #distinct(name, .keep_all = T) %>%
  drop_na(FieldID)

view(species.ds)

# all rows from Victoria's dataset have the CCA info now!
nrow(species.ds)
```

Writing the output:
``` {r output}
write.csv(species.ds, "../outputs/full.df.first.draft.csv") # add file name
```

## Merging with EUCLID's life-history dataset:

``` {r adding life-history data to env}
lifehistory <- read.csv("../data/all_traits_from_EUCLID.csv", header = T) %>%
  glimpse()
```

``` {r merge dataframes by species name}
full.ds <- left_join(species.ds, lifehistory, by = "name") %>%
  # keep one replicate per species-provenance
  distinct(UniqueID, .keep_all = T) %>%
  # remove species with no EUCLID information
  drop_na(genus)

view(full.ds)  
nrow(full.ds)
```

``` {r output}
write.csv(full.ds, "../outputs/life_history_CCA.csv")
```

We can see here that out of the 1849 taxa (rows) in the CCA dataset, 1586 have life-history trait data merged.
Let's have a look at the 263 rows that don't merge.

``` {r}
dubious_species <-
  anti_join(species.ds, lifehistory, by = "name") %>%
  distinct(name, .keep_all = T) #%>%
dubious_species <-
  dplyr::filter(!grepl( c(';?;|; x ;'), name)) %>%
  glimpse()

view(dubious_species)
nrow(dubious_species)
```

They all seem to be dubious species (either no records found for life-history traits or refered to as subspecies instead of species. but mostly no records).


## How many species from the dataset don't have life-history traits? Merging Victoria CCA species directly with euclid trait I extracted

``` {r EUCLID to Victoria species}
EUCLID <- read.csv("../data/all_traits_from_EUCLID.csv", header = T)
nrow(EUCLID)

test.species <- 
  read.csv("../data/CCA_VPM_species_PGLS.csv", header = T)
colnames(species)[4] <- "name"
test.species <-
  distinct(name, .keep_all = TRUE)
  
test.full.df <- left_join(test.species, EUCLID, by = "name")
nrow(test.full.df)
write.csv(test.full.df, "../outputs/test.full.lifehist.df.csv")
```
There are 935 species in the flora dataset and 930 in Victoria's.

## How many species in CCA don't have flora traits?
``` {r}
test.df <- left_join(EUCLID, species, by = "name")
nrow(test.df)

```
There are 930 taxa in Victoria's dataset and 823 are matching with taxa from the flora traits.


## Cleaning the dataset with Dean:

I had a discussion with Dean about the unmatched taxa in `all_traits_from_EUCLID.csv` that I extracted. After revising the taxonomy I updated the taxa names from EUCLID according to the latest classification and I am now using `EUCLIDTraitsCleaned.csv`  in order to merge it with the CCA dataset.
`EUCLIDTraitsCleaned.csv` only contains the traits that Victoria is interested in, in order to reduce the number of rows in the final dataset.

``` {r loading input data}
EUCLIDupdated <-
  read.csv("../data/EUCLIDTraitsCleaned.csv", header = T) %>%
  glimpse()

Victoria.sp <-
  read.csv("../outputs/full.df.first.draft.csv", header = T) %>%
  glimpse()

colnames(Victoria.sp)[5] <- "name2" # changing Taxonsimplified_DN to name to merge to EUCLID
colnames(Victoria.sp)[3] <- "name"
Victoria.sp[[3]] <- tolower(Victoria.sp[[3]])
glimpse(Victoria.sp) # looks good to go
```

``` {r merging datasets}
Victoria.EUCLID.df <-
    left_join(Victoria.sp, EUCLIDupdated, by = "name") %>%
  glimpse()

view(Victoria.EUCLID.df)
```

Let's have a look if there is no missing data...
``` {r writing output}
write.csv(Victoria.EUCLID.df, "../outputs/Victoria.EUCLID.df.csv")
```

There is a bunch but it's for dubious species only. Will see with Dean.

## Getting Thornhill's phylo species as csv to 

``` {r Thornhills species}
tree <- read.tree("../data/Eucalypts_ML2_dated_r8s.tre",)
thornhill.species <- as.data.frame(tree$tip.label)
view(thornhill.species)
df <- data.frame(matrix(unlist(thornhill.species), nrow=length(thornhill.species), byrow=TRUE))
write.csv(df, "../outputs/datasets_in_progress/thornhill.species.list.csv")
```

```{r cleaning thornhill phylo}
thornhill <- read.csv("../data/thornhill_species_list.csv", header = T)#%>%
view(thornhill)
  setDT(thornhill)[, paste0("taxon", 1:7) := tstrsplit(species, "_")]
write.csv(thornhill, "../outputs/split_thornhill_data.csv")
```

## How many species from Victoria list not found in Thornhill phylo? trying to merge based on species name

Thornhill's phylogeny has species names based on Dean's taxonomy so it's supposed to match perfectly with Victoria's species based on Dean's classification of Eucalypts.

``` {r matching victoria and thornhill datasets}
thornhill.species <-
  read.csv("../data/ThornhillSpeciesCleaned.csv", header = T) %>%
  glimpse()

victoria.species <-
  read.csv("../outputs/VPM_species.csv", header = T) %>%
  glimpse()

nrow(thornhill.species)
nrow(victoria.species) # there are 200 taxa more in Victoria's compared to Thornhill

thornhill.vic.df <- left_join(victoria.species, thornhill.species, by = "name") %>%
  #removing rows with no phylo information
  drop_na(original)

nrow(thornhill.vic.df) # looks like not all Thornhill species have a match
                       # (only 603 compared to Thornhill = 729)
view(thornhill.vic.df)
write.csv(thornhill.vic.df, "../outputs/datasets_in_progress/Victoria_thornhill_merge.csv")
```

``` {r adding Thornhill voucher data to Victoria's dataset}
thornhill.species <-
  read.csv("../data/ThornhillDataCleaned.csv", header = T) %>%
  glimpse()
colnames(thornhill.species)[1] <- "name"
thornhill.species[[1]] <- tolower(thornhill.species[[1]])
glimpse(thornhill.species)
Victoria.EUCLID.df <-
    read.csv("../outputs/Victoria.EUCLID.df.csv", header = T) %>%
  glimpse()


FULL.DF <-
  left_join(Victoria.EUCLID.df, thornhill.species, by = "name") %>%
  distinct(UniqueID, .keep_all = TRUE) %>%
  glimpse()

# I need to add the original name in EUCLID
EUCLID.traits.df <- read.csv("../data/all_traits_from_EUCLID.csv", header = T)
colnames(EUCLID.traits.df)[1] <- "EUCLID_name"
glimpse(EUCLID.traits.df)

full.FULL.DF <-
  left_join(Victoria.EUCLID.df, EUCLID.traits.df, by = "botanical_desc") %>%
  distinct(UniqueID, .keep_all = TRUE)
```

``` {r}
write.csv(FULL.DF, "../outputs/datasets_in_progress/Victoria_EUCLID_Thornhill_merge.v2.csv")
```
``` {r}
write.csv(full.FULL.DF, "../outputs/datasets_in_progress/Victoria_EUCLID_Thornhill_merge.v3.csv")
```

# Conclusion: not all Victoria's taxa are listed in Thornhill and same for the opposite, not all Thornhill's taxa are in Victoria's. Total, 603 taxa are matching between the two.

If everything is fine, the only data missing now in the merged dataset corresponds to taxa that Victoria has but were not included in the phylogeny (so we'll have to suss out the phylo signal for those ones)

Now let's try to add the functional trait to the back end of that, and depending were the gaps are across the three datasets we'll know if it's a lack of data because EUCLID simply doesn't have data for those taxa or if it's dubious taxa and a species in EUCLID is actually a subspecies in Thornhill/DN.

```{r second merging with life-history traits from EUCLID}
Victoria.thornhill.merge.df <- read.csv("../outputs/datasets_in_progress/Victoria_thornhill_merge.csv", header = T)

nrow(Victoria.thornhill.merge.df)
view(Victoria.thornhill.merge.df)
EUCLID.traits.df <- read.csv("../data/all_traits_from_EUCLID.csv", header = T)
view(EUCLID.traits.df)
nrow(EUCLID.traits.df)

Victoria_EUCLID_merge <- left_join(victoria.species, EUCLID.traits.df, by = "name")
# now adding a column that says if present in Thornhill taxa list or not:
Victoria_EUCLID_merge$in.thornhill <- victoria.species$name %in% thornhill.species$name

view(Victoria_EUCLID_merge)
nrow(Victoria_EUCLID_merge)
```

``` {r output}
write.csv(Victoria_EUCLID_merge, "../outputs/datasets_in_progress/victoria_thornhill_euclid_merge.csv")
```

We now have a dataset with the life-history trait data for Eucalypt species and we know which ones are present in the taxa list from Thornhill's phylogeny.


## Final dataset:

I cleaned the EUCLID phylogeny and added an additional column with the updated taxon names from the classification.
Now all I need to do is merge Victoria's dataset to EUCLID, add the resprouting data from Dean, match the taxa to Thornhill's voucher data.

``` {r final merge}
#### input data###

Victoria <-
  read.csv("../data/CCA_VPM_species_PGLS.csv", header = T) %>%
  distinct(UniqueID, .keep_all = TRUE) %>%
  glimpse()
colnames(Victoria)[2] <- "name" # by updated name
colnames(Victoria)[5] <- "name" # by ALA name

glimpse(Victoria)

EUCLID <-
  read.csv("../data/EUCLIDTraitsCleaned.csv", header = T) %>%
  glimpse()
colnames(EUCLID)[2] <- "name"
glimpse(EUCLID)

Resprouting <-
  read.csv("../data/CCARegenerationStrategyCleaned.csv", header = T) %>%
  glimpse()
colnames(Resprouting)[12] <- "name"
glimpse(Resprouting)

Thornhill <-
  read.csv("../data/ThornhillDataCleaned.csv", header = T) %>%
  glimpse()
colnames(Thornhill)[2] <- "name"
glimpse(Thornhill)


# merge Victoria to Resprouting by name
Victoria.Resp <-
  left_join(Victoria, Resprouting, by = "name") %>%
  glimpse()

# merge Victoria.Resp to EUCLID by DN.name = name
Victoria.Resp.EUCLID <-
  left_join(Victoria.Resp, EUCLID, by = "name") %>%
  glimpse()

# merge Victoria.Resp.EUCLID to Thornhill by name = binomial
colnames(Victoria.Resp.EUCLID)[2] <- "name2" # by updated name
colnames(Victoria.Resp.EUCLID)[5] <- "name" # by ALA name
glimpse(Victoria.Resp.EUCLID)

Victoria.Resp.EUCLID.Thornhill <-
  left_join(Victoria.Resp.EUCLID, Thornhill, by = "name") %>%
  distinct(UniqueID, .keep_all = TRUE) %>%
  glimpse()

# write resulting df
write.csv(Victoria.Resp.EUCLID.Thornhill, "../outputs/03092021.VPM.PGLS.df.csv")
# anti_join Victoria.Resp.EUCLID to Thornhill
# (check which taxa don't merge)
non.matching.Thornhill.species <-
  anti_join(Thornhill, Victoria, by = "name")
# write resulting df
write.csv(non.matching.Thornhill.species, "../outputs/NonMatchingThornhillSpecies.csv")

# Thornhill uses the outaded names too so merge with EUCLID.name instead of updated name

Victoria.Thornhill.by.ALA <-
  left_join(Victoria, Thornhill, by = "name") %>%
  distinct(UniqueID, .keep_all = TRUE) %>%
  glimpse()
write.csv(Victoria.Thornhill.by.ALA, "../outputs/Victoria.Thornhill.by.ALA.csv")

```

## Making the phylogeny's tip labels match with victoria's dataset:

``` {r changing phylo tip labels}
thornhill_labels <- read.csv("../data/tree_labels.csv")
tree <- read.tree("../data/Eucalypts_ML2_dated_r8s.tre")

tree$tip.label<-thornhill_labels$name3[match(tree$tip.label, thornhill_labels$original)] # changing tip label to match victoria species name
tree$tip.label<-sapply(tree$tip.label, function(x) parse(text=x))
ggtree(tree)
plot(tree)
ggsave("../plots/test.pdf", width = 50, height = 120, limitsize = FALSE)
```


