---
title: "Constructing PGLS dataset for CCA species with functional and life-history trait data"
author: "Anne-Cecile Colin"
date: "10/06/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(devtools)
library(readr)
library(sp)
library(raster)
library(stats)
library(plyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(conflicted)
library(ape)
```

## Merging life-history trait data to CCA# by = species

``` {r adding life-history data}

CCAnumber <- read.csv("../data/life_history_CCA.csv", header = T)
view(CCAnumber)
nrow(CCAnumber)

species <- read.csv("../data/CCA_VPM_species_PGLS.csv", header = T) %>%
  distinct(acceptedname_ala, .keep_all = TRUE)
species <- species[!(species$Taxonresolution_ALA=="Hybrid") | species$Taxonresolution_ALA=="Variety",]
colnames(species)[4] <- "name"
nrow(species)
view(species)
write.csv(species, "../outputs/VPM_species.csv") # just to check the number of species in each genus

trait.ds <- full_join(species, CCAnumber, by = "name") %>%
  distinct(name, .keep_all = T) %>%
  drop_na(FieldID)

view(trait.ds)
nrow(trait.ds)

write.csv(trait.ds, "../outputs/full.df.first.draft.csv") # add file name
```

## How many species from the dataset don't have life-history traits? Merging Victoria CCA species directly with euclid trait from Desi

``` {r EUCLID to Victoria species}
EUCLID <- read.csv("../data/euc_traits_from_EUCLID.csv", header = T)
view(EUCLID)
nrow(EUCLID)

test.full.df <- left_join(species, EUCLID, by = "name")
nrow(test.full.df)
write.csv(test.full.df, "../outputs/test.full.lifehist.df.csv")
```
There are 823 species in the flora dataset and 930 in Victoria's.
Corymbia and Angophora species have no trait data (96), 

## How many species in CCA don't have flora traits?
``` {r}
test.df <- left_join(EUCLID, species, by = "name")
nrow(test.df)

```
There are 930 taxa in Victoria's dataset and 823 are matching with taxa from the flora traits.



## Getting Thornhill's phylo species as csv to 

``` {r Thornhills species}
tree <- read.tree("../data/Eucalypts_ML2_dated_r8s.tre",)
thornhill.species <- as.data.frame(tree$tip.label)
view(thornhill.species)
df <- data.frame(matrix(unlist(thornhill.species), nrow=length(thornhill.species), byrow=TRUE))
write.csv(df, "../outputs/thornhill.species.list.csv")

thornhill <- read.csv("../data/thornhill_species_list.csv")
nrow(thornhill)
```

## How many species from Victoria list not found in Thornhill phylo?
``` {r}
```



