---
title: "Cleaning the dubious taxonomy from the PGLS dataset"
author: "Anne-Cecile Colin"
date: "10/06/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

librarian::shelf(tidyverse, readr, conflicted, ggtree, ape, data.table, DesiQuintans/desiderata)

```

## 1. Aim:

In the previous script `1_constructing_PGLS_datasets.Rmd`, I merged and cleaned the life-history traits from EUCLID with Victoria's CCA dataset.

The aim of this script is to use the datasets created from `1_constructing_PGLS_datasets.Rmd` to figure out the best match for taxa names across the three datasets: the list of taxa from Thornhill's phylogeny, the taxa at CCA that Victoria uses for her analysis, and the taxa specified in the EUCLID flora.

The descriptions in the EUCLID flora can be pretty old and therefore outdated in terms of taxonomy (what names the taxa are refered to as).
With this script I will make a list of the unmatched taxa across the three datasets and update the taxa names in my EUCLID dataset so they can all match to their current taxonomical name.

At the end of this script I will have a different lists of unmatched taxa among the three datasets. I will use those lists to count the number of dubious species in the EUCLID dataset with Dean, and change the taxa name according to the classification of Eucalypts.

## 2. Read input data

``` {r}
EUCLID <-
  read.csv("../data/all_traits_from_EUCLID.csv", header = T) %>%
  glimpse()

Victoria_full_df <-
  read.csv("../outputs/datasets_in_progress/Victoria_EUCLID_merge.csv", header = T) %>%
  distinct(name, .keep_all = T) %>%
  glimpse()

view(Victoria_full_df)
```

# Which taxa (and how many) are in the EUCLID flora but not in CCA?

The function `dyplr::anti_join` retunrs rows of x that do not have a match in y.

``` {r}
CCA.taxa.not.in.EUCLID <-
  anti_join(Victoria_full_df, EUCLID, by = "name")

nrow(CCA.taxa.not.in.EUCLID) # 137
view(CCA.taxa.not.in.EUCLID)
```

137 taxa in Victoria's dataset do not have a match in EUCLID.

``` {r}
EUCLID.taxa.not.in.CCA <-
  anti_join(EUCLID, Victoria_full_df, by = "name")
nrow(EUCLID.taxa.not.in.CCA) # 160
View(EUCLID.taxa.not.in.CCA)
```

I'm going to manually check all of those taxa to see if they are really present in CCA or just that the name is different.

Then I'll make a column with the name it should be refered to as to match with CCA.
``` {r}
write.csv(EUCLID.taxa.not.in.CCA, "../outputs/species.to.check.with.Dean.csv")
```
160 taxa in EUCLID do not have a match in Victoria's dataset.
Most of them have a different taxonomy where it is a subsp. in one and a species in the other.
Let's try to match them with the ALA_accepted name to see if the list reduces.

# Trying to match the remaining taxa in Victoria's dataset with the ALA taxa name

``` {r}
second.merge.by.name <-
  left_join(CCA.taxa.not.in.EUCLID, EUCLID.taxa.not.in.CCA, BY = C("name2" = "name"))
second.merge.by.name <- na.omit(second.merge.by.name, cols=Row)

glimpse(second.merge.by.name)

nrow(second.merge.by.name)
view(second.merge.by.name)
```

