---
title: "Manipulating phylo and plotting data"
author: "AC"
date: "16/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
librarian::shelf(dplyr, ape, conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("drop.tip", "ape")
librarian::shelf(tidyverse, ggtree, phytools, TreeTools, tidytree, treeio)

```

```{r, raw data}
thornhill_labels <- read.csv("data/tree_labels.csv")
tree <- read.tree("data/Eucalypts_ML2_dated_r8s.tre")
```


## First tree manipulation: changing the tip labels

Let's have a look at what they look like now:
```{r}
ggtree(tree,layout = "fan")
```

In the environment, "tree" is a phylo object and if you click on it you can see tip.label as a list (type `tree[["tip.label"]]` in the console) which gives the original name of the tree's leaves contained in Thornhill's phylogeny.  

Now we want to swap those names with the ones we are going to use (the CCA names) based on the spreadsheet that we edited with Dean.  

```{r}
new_tree <- tree
new_tree$tip.label <- thornhill_labels$name5[TreeTools::match(tree$tip.label, thornhill_labels$original)]
par(mfrow=c(1,2))
plot(tree)
plot(new_tree)
```
I'm using the match() function of TreeTools to change the tip.label. As you can see in the phylo object new_tree, tip.label is still read as a character just like in the original tree, and the names appear normally when plotted. Nothing else should differ between the two phylo objects.  
Note to myself: there are other ways to change the tip labels but this set of commands is the only way to conserve tip.label as character, ensuring that the phylo object remains readable by phylo packages in the following steps. If tip.label is not displayed as character then any manipulation of tips will fail as the function is unable to retrieve them in the phylo object.  

If you open thornhill_labels, you will see that the variable "name3" corresponds to the full species name with underscores instead of white spaces as they can be problematic for some manipulations with phylogenetic packages in R.  
'name5' corresponds to the binomial with underscores as white spaces.
The names will match between the binomial in Thornhill, and each taxa name in CCA.

## Second tree manipulation: removing some tips

Now we want to remove species in the phylogeny that are not part of the analysis so they do not appear in the modified phylogeny. To do so I use the function drop.tip() from TreeTools, which removes a set of tips and collapses to the corresponding set of branches.

```{r}
pruned_tree <- new_tree

prune.species <- c("lophostemon_suaveolens", "myrtus_communis", "decaspermum_humile",
                   "amomyrtus_luma", "cloezia_floribunda", "tepualia_stipularis",
                   "acmena_smithii", "syzygium_angophoroides", "syncarpia_hillii",
                   "syncarpia_glomulifera", "hypocalymma_linifolium", "kunzea_ericoides",
                   "agonis_flexuosa", "allosyncarpia_ternata", "stockwellia_quadrifida",
                   "stockwellia_quadrifida", "eucalyptopsis_papuana", "eucalyptopsis_alauda",
                   "arillastrum_gummiferum", "lophostemon_confertus", "kjellbergiodendron_celebicum",
                   "heteropyxis_natalensis", "eucalyptus_yangoura", "corymbia_novoguinensis",
                   "eucalyptus_wetarensis", "eucalyptus_websteriana", "eucalyptus_urophylla",
                   "corymbia_paractia", "corymbia_setosa", "eucalyptus_arachnaeae",
                   "eucalyptus_baiophylla", "eucalyptus_coccifera", "eucalyptus_glomericassis",
                   "eucalyptus_grasbyi", "eucalyptus_leucophylla", "eucalyptus_stockeri",
                   "eucalyptus_sheathiana", "kjellbergiodendron_celebicum", "lophostemon_confertus")

pruned_tree <- drop.tip(pruned_tree, prune.species, trim.internal = T)
plot(pruned_tree)

#write.tree(pruned_tree, "../plots/tree_inprogress_check_taxanames_in_double.tre")
```

You can see at the base of the tree that species outside of Angophora, Corymbia and Eucalyptus are gone and the tree collapsed.  

If you look at the phylo object 'pruned_tree' you can see that the listed tips in the vector 'prune.species' are removed from tip.label and that the number of tips and nodes is now reduced.  

## Third tree manipulation: adding tips

Now we can change the tree to include the additional tips for all subspecies, hybrids and provenances at CCA.

First, we need to create a matrix with the required data for the function bind.tip() to use.
I transform the tree as a tibble to extract the information I require from the existing tips (tips' labels and node numbers) to use the bind.tip() function to add the new tips.

I only kept the information that bind.tip requires, the node number to which the new tip will be attached and its corresponding new label.
``` {r}
all_tree <- pruned_tree
new_tips_data <- read.csv("data/NewTips(onlyduplicates).csv") # only duplicates = only contains the tips from the current tree that will be receiving new tips i.e. taxonomically closest species absent in Thornhill that need a tip, all current placements on the phylogeny in which trees at CCA will be put. not the tips that will remain unchanged (which I manually removed from the dataframe). Unique tips (~30/50 tips from what I recall) include tips that will receive only one uniqueID

tibble_tree <- as_tibble(all_tree) %>% # putting the full_tree$tips and
  as_data_frame() %>%                   # full_tree$branchlength to the right format to merge
  glimpse()

branch_length_matrix <- left_join(new_tips_data, tibble_tree, by = c("phylo.tip.label.match" = "label")) %>% # merging
  as_data_frame() %>%
  select("uniqueID", "node") %>% # keeping only the important stuff
  as.matrix() %>% 
  glimpse()

#write_csv(branch_length_matrix, "bind.tip_tip_matrix.csv") # saved it just in case


head(all_tree)
head(branch_length_matrix)
?bind.tip


```

To use bind.tip() on all tips I then made a for loop on the function.
Now let's add all the new tips for each CCA taxa to the tree.
Matrix providing data for each new tip:
  - column 1: the label of the new tip
  - column 2: the node number to which this tip will be anchored
  - column 3: the branch length for this new tip

```{r}
bldf <- as.data.frame(branch_length_matrix)
bldf$node <- as.numeric(bldf[,2])
head(bldf)


test_tree2 <- all_tree

for (i in 1:length(bldf$uniqueID)) { # the for loop takes a minute to run
  labs <- bldf$uniqueID[i]
  nodes <- which(test_tree2$tip.label == all_tree$tip.label[bldf$node[i]])
  test_tree2 <- bind.tip(test_tree2, tip.label = labs, edge.length= 0, where=nodes)
}

plot(test_tree2, cex = 0.25)

write.tree(test_tree2, "plots/with_new_tips_as_unique_ID_full_tree.tre")

```


## Final tree manipulation:

We can't really see it in the plots in Rstudio but the tree still contains the original tree labels 
(without the unique ID) where the new tips have been anchored.
For example on node 10 for Angophora costata we now have:
  `2073`
  `6796`
  `2082`
  `3206`
  and the original `angophora_costata`
Same goes on for all of them.

We want to:
(1) change the labels for the tips that didn't need to have new tips added
(2) prune the tree a final time to only have the labels with the unique IDs

```{r}

pruned_tree2 <-read.tree("plots/with_new_tips_as_unique_ID_full_tree.tre")

prune.species <- c("angophora_costata", "angophora_hispida", "corymbia_arenaria", "corymbia_brachycarpa",
                  "corymbia_cadophora", "corymbia_calophylla", "corymbia_dichromophloia", "corymbia_papillosa",
                  "corymbia_chippendalei", "corymbia_haematoxylon", "corymbia_citriodora", "corymbia_erythrophloia",
                  "corymbia_greeniana", "corymbia_deserticola", "corymbia_ellipsoidea", "corymbia_eremaea",
                  "corymbia_ferruginea", "corymbia_ficifolia", "corymbia_grandifolia", "corymbia_hamersleyana",
                  "corymbia_lenziana", "corymbia_ligans", "corymbia_henryi", "corymbia_nesophila", "corymbia_opaca",
                  "corymbia_pachycarpa", "corymbia_polycarpa", "corymbia_polysciada", "corymbia_ptychocarpa",
                  "corymbia_stockeri", "corymbia_terminalis", "corymbia_trachyphloia", "corymbia_watsoniana",
                  "eucalyptus_acaciiformis", "eucalyptus_acmenoides", "eucalyptus_decipiens", "eucalyptus_viridis",
                  "eucalyptus_aequioperta", "eucalyptus_trivalvis", "eucalyptus_umbra", "eucalyptus_agglomerata",
                  "eucalyptus_aggregata", "eucalyptus_cypellocarpa", "eucalyptus_kingsmillii", "eucalyptus_alba",
                  "eucalyptus_albens", "eucalyptus_albida", "eucalyptus_albopurpurea", "eucalyptus_alipes",
                  "eucalyptus_alligatrix", "eucalyptus_nudicaulis", "eucalyptus_amplifolia", "eucalyptus_amygdalina",
                  "eucalyptus_andrewsii", "eucalyptus_angulosa", "eucalyptus_angustissima", "eucalyptus_kessellii",
                  "eucalyptus_apiculata", "eucalyptus_apothalassica", "eucalyptus_approximans", "eucalyptus_aquilina",
                  "eucalyptus_arachnaea", "eucalyptus_arborella", "eucalyptus_aromaphloia", "eucalyptus_gunnii",
                  "eucalyptus_arenacea", "eucalyptus_argutifolia", "eucalyptus_lucasii", "eucalyptus_erythronema",
                  "eucalyptus_aspersa", "eucalyptus_aspratilis", "eucalyptus_assimilans", "eucalyptus_astringens",
                  "eucalyptus_miniata", "eucalyptus_baeuerlenii", "eucalyptus_baileyana", "eucalyptus_bakeri",
                  "eucalyptus_balladoniensis", "eucalyptus_interstans", "eucalyptus_goniocalyx", "eucalyptus_barberi",
                  "eucalyptus_ovularis", "eucalyptus_baueriana", "eucalyptus_baxteri", "eucalyptus_beardiana",
                  "eucalyptus_behriana", "eucalyptus_bensonii", "eucalyptus_benthamii", "eucalyptus_globulus",
                  "eucalyptus_deglupta", "eucalyptus_pellita", "eucalyptus_blakelyi", "eucalyptus_blaxlandii",
                  "eucalyptus_bosistoana", "eucalyptus_botryoides", "eucalyptus_brachycalyx", "eucalyptus_tetraptera",
                  "eucalyptus_brassiana", "eucalyptus_brevifolia", "eucalyptus_bridgesiana", "eucalyptus_brockwayi",
                  "eucalyptus_brookeriana", "eucalyptus_hallii", "eucalyptus_buprestium", "eucalyptus_caesia",
                  "eucalyptus_calcareana", "eucalyptus_calcicola", "eucalyptus_caleyi", "eucalyptus_caliginosa",
                  "eucalyptus_calycogona", "eucalyptus_camaldulensis", "eucalyptus_largeana", "eucalyptus_mckieana",
                  "eucalyptus_campaspe", "eucalyptus_camphora", "eucalyptus_canescens", "eucalyptus_capillosa",
                  "eucalyptus_ceratocorys", "eucalyptus_capitellata", "eucalyptus_captiosa", "eucalyptus_carnea",
                  "eucalyptus_carnei", "eucalyptus_celastroides", "eucalyptus_cernua", "eucalyptus_dalrympleana",
                  "eucalyptus_chartaboma", "eucalyptus_chlorophylla", "eucalyptus_cinerea", "eucalyptus_cladocalyx",
                  "eucalyptus_clelandii", "eucalyptus_clivicola", "eucalyptus_comitae-vallis", "eucalyptus_concinna",
                  "eucalyptus_conferruminata", "eucalyptus_confluens", "eucalyptus_conglobata", "eucalyptus_conspicua",
                  "eucalyptus_conveniens", "eucalyptus_coolabah", "eucalyptus_cooperiana", "eucalyptus_cordata",
                  "eucalyptus_cornuta", "eucalyptus_corrugata", "eucalyptus_corticosa", "eucalyptus_melanoleuca",
                  "eucalyptus_cosmophylla", "eucalyptus_crebra", "eucalyptus_creta", "eucalyptus_cretata",
                  "eucalyptus_crucis", "eucalyptus_exilipes", "eucalyptus_cunninghamii", "eucalyptus_cupularis",
                  "eucalyptus_cylindrocarpa", "eucalyptus_magnificata", "eucalyptus_dealbata", "eucalyptus_decurva",
                  "eucalyptus_deflexa", "eucalyptus_delegatensis", "eucalyptus_delicata", "eucalyptus_burgessiana",
                  "eucalyptus_densa", "eucalyptus_dielsii", "eucalyptus_diminuta", "eucalyptus_discreta",
                  "eucalyptus_uncinata", "eucalyptus_dissimulata", "eucalyptus_distans", "eucalyptus_distuberosa",
                  "eucalyptus_diversicolor", "eucalyptus_diversifolia", "eucalyptus_dives", "eucalyptus_dolichocera",
                  "eucalyptus_falcata", "eucalyptus_drummondii", "eucalyptus_dumosa", "eucalyptus_dundasii",
                  "eucalyptus_dura", "eucalyptus_dwyeri", "eucalyptus_ebbanoensis", "eucalyptus_petrensis",
                  "eucalyptus_educta", "eucalyptus_effusa", "eucalyptus_elaeophloia", "eucalyptus_elata",
                  "eucalyptus_elliptica", "eucalyptus_erectifolia", "eucalyptus_eremicola", "eucalyptus_eremophila",
                  "eucalyptus_eugenioides", "eucalyptus_erythrocorys", "eucalyptus_eudesmioides",
                  "eucalyptus_ewartiana", "eucalyptus_exigua", "eucalyptus_exilis", "eucalyptus_extensa",
                  "eucalyptus_willisii", "eucalyptus_fasciculosa", "eucalyptus_paniculata", "eucalyptus_fibrosa",
                  "eucalyptus_histophylla", "eucalyptus_flindersii", "eucalyptus_flocktoniae", "eucalyptus_foecunda",
                  "eucalyptus_formanii", "eucalyptus_forrestiana", "eucalyptus_fraseri", "eucalyptus_fraxinoides",
                  "eucalyptus_frenchiana", "eucalyptus_froggattii", "eucalyptus_fruticosa", "eucalyptus_gamophylla",
                  "eucalyptus_gardneri", "eucalyptus_georgei", "eucalyptus_gillenii", "eucalyptus_gillii",
                  "eucalyptus_gittinsii", "eucalyptus_glaucescens", "eucalyptus_glaucina", "eucalyptus_globoidea",
                  "eucalyptus_gongylocarpa", "eucalyptus_goniantha", "eucalyptus_gracilis", "eucalyptus_grandis",
                  "eucalyptus_griffithsii", "eucalyptus_grisea", "eucalyptus_grossa", "eucalyptus_guilfoylei",
                  "eucalyptus_gypsophila", "eucalyptus_hebetifolia", "eucalyptus_herbertiana", "eucalyptus_horistes",
                  "eucalyptus_hypolaena", "eucalyptus_incerata", "eucalyptus_incrassata", "eucalyptus_indurata",
                  "eucalyptus_infera", "eucalyptus_lockyeri", "eucalyptus_insularis", "eucalyptus_intertexta",
                  "eucalyptus_jacksonii", "eucalyptus_jimberlanica", "eucalyptus_subcrenulata", "eucalyptus_jucunda",
                  "eucalyptus_jutsonii", "eucalyptus_tereticornis", "eucalyptus_kitsoniana", "eucalyptus_kochii",
                  "eucalyptus_kondininensis", "eucalyptus_kruseana", "eucalyptus_kumarlensis", "eucalyptus_lacrimans",
                  "eucalyptus_melanoxylon", "eucalyptus_lanepoolei", "eucalyptus_langleyi", "eucalyptus_lansdowneana",
                  "eucalyptus_largiflorens", "eucalyptus_lehmannii", "eucalyptus_leptocalyx", "eucalyptus_leptophleba",
                  "eucalyptus_leptophylla", "eucalyptus_leptopoda", "eucalyptus_lesouefii", "eucalyptus_leucophloia",
                  "eucalyptus_tephrodes", "eucalyptus_leucoxylon", "eucalyptus_ligulata", "eucalyptus_litorea",
                  "eucalyptus_longicornis", "eucalyptus_longifolia", "eucalyptus_loxophleba", "eucalyptus_luculenta",
                  "eucalyptus_luehmanniana", "eucalyptus_macarthurii", "eucalyptus_macrandra", "eucalyptus_macrocarpa",
                  "eucalyptus_macrorhyncha", "eucalyptus_scias", "eucalyptus_major", "eucalyptus_mannensis",
                  "eucalyptus_mannifera", "eucalyptus_marginata", "eucalyptus_mcquoidii", "eucalyptus_megacarpa",
                  "eucalyptus_megacornuta", "eucalyptus_melanophloia", "eucalyptus_melliodora", "eucalyptus_merrickiae",
                  "eucalyptus_micranthera", "eucalyptus_microcarpa", "eucalyptus_mimica", "eucalyptus_minniritchi",
                  "eucalyptus_moderata", "eucalyptus_moluccana", "eucalyptus_moorei", "eucalyptus_morrisbyi",
                  "eucalyptus_morrisii", "eucalyptus_multicaulis", "eucalyptus_myriadena", "eucalyptus_nandewarica",
                  "eucalyptus_nitida", "eucalyptus_neglecta", "eucalyptus_neutra", "eucalyptus_nicholii",
                  "eucalyptus_nobilis", "eucalyptus_normantonensis", "eucalyptus_nortonii", "eucalyptus_nutans",
                  "eucalyptus_obconica", "eucalyptus_obesa", "eucalyptus_obliqua", "eucalyptus_obtusiflora",
                  "eucalyptus_occidentalis", "eucalyptus_polybractea", "eucalyptus_oldfieldii", "eucalyptus_oleosa",
                  "eucalyptus_oligantha", "eucalyptus_olivina", "eucalyptus_olsenii", "eucalyptus_optima",
                  "eucalyptus_oraria", "eucalyptus_orbifolia", "eucalyptus_oreades", "eucalyptus_ornata",
                  "eucalyptus_orthostemon", "eucalyptus_ovata", "eucalyptus_pachyloma", "eucalyptus_pachyphylla",
                  "eucalyptus_paedoglauca", "eucalyptus_panda", "eucalyptus_mooreana", "eucalyptus_parramattensis",
                  "eucalyptus_patens", "eucalyptus_pauciflora", "eucalyptus_perangusta", "eucalyptus_perriniana",
                  "eucalyptus_petiolaris", "eucalyptus_petraea", "eucalyptus_phaenophylla", "eucalyptus_phenax",
                  "eucalyptus_phoenicea", "eucalyptus_pileata", "eucalyptus_pimpiniana", "eucalyptus_piperita",
                  "eucalyptus_platycorys", "eucalyptus_platyphylla", "eucalyptus_pleurocarpa", "eucalyptus_polita",
                  "eucalyptus_polyanthemos", "eucalyptus_preissiana", "eucalyptus_prolixa", "eucalyptus_youmanii",
                  "eucalyptus_propinqua", "eucalyptus_protensa", "eucalyptus_pruinosa", "eucalyptus_pterocarpa",
                  "eucalyptus_pulverulenta", "eucalyptus_purpurata", "eucalyptus_quadrans", "eucalyptus_quadricostata",
                  "eucalyptus_quaerenda", "eucalyptus_racemosa", "eucalyptus_radiata", "eucalyptus_rameliana",
                  "eucalyptus_ravida", "eucalyptus_recta", "eucalyptus_tindaliae", "eucalyptus_pluricaulis",
                  "eucalyptus_regnans", "eucalyptus_resinifera", "eucalyptus_rigens", "eucalyptus_rigidula",
                  "eucalyptus_rodwayi", "eucalyptus_rosacea", "eucalyptus_roycei", "eucalyptus_rubida",
                  "eucalyptus_rudderi", "eucalyptus_rudis", "eucalyptus_rugosa", "eucalyptus_rugulata",
                  "eucalyptus_salicola", "eucalyptus_saligna", "eucalyptus_salmonophloia", "eucalyptus_salubris",
                  "eucalyptus_sargentii", "eucalyptus_saxatilis", "eucalyptus_scoparia", "eucalyptus_surgens",
                  "eucalyptus_sepulcralis", "eucalyptus_serraensis", "eucalyptus_vittata", "eucalyptus_sideroxylon",
                  "eucalyptus_singularis", "eucalyptus_smithii", "eucalyptus_socialis", "eucalyptus_urna",
                  "eucalyptus_nerthicola", "eucalyptus_sparsa", "eucalyptus_sparsifolia", "eucalyptus_spathulata",
                  "eucalyptus_stricta", "eucalyptus_spreta", "eucalyptus_steedmanii", "eucalyptus_stellulata",
                  "eucalyptus_stricklandii", "eucalyptus_strzeleckii", "eucalyptus_subangusta", "eucalyptus_subtilis",
                  "eucalyptus_suggrandis", "eucalyptus_synandra", "eucalyptus_talyuberlup", "eucalyptus_tholiformis",
                  "eucalyptus_tenera", "eucalyptus_tenuipes", "eucalyptus_tenuiramis", "eucalyptus_terebra",
                  "eucalyptus_thamnoides", "eucalyptus_todtiana", "eucalyptus_tricarpa", "eucalyptus_triflora",
                  "eucalyptus_ultima", "eucalyptus_utilis", "eucalyptus_redunca", "eucalyptus_vegrandis",
                  "eucalyptus_vernicosa", "eucalyptus_verrucata", "eucalyptus_vicina", "eucalyptus_victoriana",
                  "eucalyptus_victrix", "eucalyptus_viminalis", "eucalyptus_vokesensis", "eucalyptus_retinens",
                  "eucalyptus_wandoo", "eucalyptus_woodwardii", "eucalyptus_wyolensis", "eucalyptus_impensa",
                  "eucalyptus_lateritica", "eucalyptus_xanthonema", "eucalyptus_xerothermica", "eucalyptus_yalatensis",
                  "eucalyptus_yilgarnensis", "eucalyptus_youngiana", "eucalyptus_yumbarrana")


pruned_tree2 <- drop.tip(pruned_tree2, prune.species, trim.internal = T)

tree_tipsID <- pruned_tree2
plot(pruned_tree2)

second.prunning <- c("angophora_bakeri", "angophora_exul", "angophora_robur", "angophora_woodsiana", 
                     "corymbia_aparrerinja", "corymbia_arafurica", "corymbia_blakei", "corymbia_candida",
                     "corymbia_catenaria", "corymbia_confertiflora", "corymbia_dallachiana", "corymbia_dampieri",
                     "corymbia_dendromerinx", "corymbia_dunlopiana", "corymbia_flavescens", "corymbia_gilbertensis",
                     "corymbia_hylandii", "corymbia_lamprophylla", "corymbia_novaguineesis", "corymbia_opacula",
                     "corymbia_papuana", "corymbia_pauciseta", "corymbia_porphyritica", "corymbia_tessellaris",
                     "corymbia_torelliana", "corymbia_torta", "corymbia_variegata", "eucalyptus_acroleuca",
                     "eucalyptus_alaticaulis", "eucalyptus_argyphea", "eucalyptus_brachycorys", "eucalyptus_carnabyi",
                     "eucalyptus_communalis", "eucalyptus_cuspidata", "eucalyptus_disclusa", "eucalyptus_dolorosa",
                     "eucalyptus_fergusonii", "eucalyptus_ignorabilis", "eucalyptus_imlayensis", "eucalyptus_limitaris",
                     "eucalyptus_medialis", "eucalyptus_microschema", "eucalyptus_minigwalica", "eucalyptus_placita",
                     "eucalyptus_plauta", "eucalyptus_ralla", "eucalyptus_scyphocalyx", "eucalyptus_tectifica",
                     "eucalyptus_tintinnans")

pruned_tree2 <- drop.tip(pruned_tree2, second.prunning, trim.internal = T)
plot(pruned_tree2)
write.tree(pruned_tree2, "plots/check-prunning_full_tree.tre")

tree_tipsID <- pruned_tree2

```


Just checking what's left to do on the tree...
```{r}

tibble_tree2 <- as_tibble(tree_tipsID$tip.label) %>% # putting the full_tree$tips and
  as_data_frame() %>%                   # full_tree$branchlength to the right format to merge
  glimpse()
write_csv(tibble_tree2, "checking_species_names_tree_matrix.csv") # compare list of species names and their numbers
```
I checked and everything that is on the tree will be renamed.

```{r}
ID_tips_data <- read.csv("data/rename.fulltree.with.uniqueID.csv")

tree_tipsID$tip.label <- ID_tips_data$new[TreeTools::match(tree_tipsID$tip.label, ID_tips_data$old)]
par(mfrow=c(1,2))
plot(tree_tipsID)

write.tree(tree_tipsID, "plots/final_tree_uniqueID.tre") ######### !! FINALLY !! ##########
```
Now we have the final tree to map the traits!!!
