---
title: "Plotting_phylo_tree"
author: "AC"
date: "04/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
librarian::shelf(tidyverse, readr, conflicted, ggtree, ape, data.table, phytools, TreeTools,tidytree,  DesiQuintans/desiderata)
conflict_prefer("filter", "dplyr")
```

## Plotting Thornhill phylogeny

Now that we have that damn tree we can plot it and map traits onto it!

```{r}
tree <- read.tree("plots/final_tree_uniqueID.tre")
info <- read.csv("data/DroughtImpact_SummaryBy_Completetraits.csv")
```

Just visualizing the data beforehand
```{r}
plot(tree, "fan")
view(info)
```


Let's start with mapping the drought incidence
```{r}
p1 <- ggtree(tree, layout='circular') %<+% info + 
  geom_tippoint(aes(color=Drought2020Scale.mean)) +
  geom_tile(aes(color=Drought2020Scale.mean)) +
  scale_color_gradientn(colours=c("blue", "green", "yellow", "red")) +
  geom_tiplab(aes(color=Drought2020Scale.mean, ftype="i",fsize=0.5), hjust = -.1)
print(p1)
ggsave("plots/Drought2020Scale.mapped.colours.pdf", width = 50, height = 120, limitsize = FALSE)

p2 <- ggtree(tree, layout='circular') %<+% info + 
  geom_tippoint(aes(color=AICorrected)) +
  scale_color_gradientn(colours=c("blue", "green", "yellow", "red")) +
  geom_tiplab(aes(color=AICorrected, ftype="i",fsize=0.5), hjust = -.1)
print(p2)
ggsave("plots/AICorrected.mapped.colours.pdf", width = 50, height = 120, limitsize = FALSE)

p3 <- ggtree(tree, layout='rectangular') %<+% info + 
  geom_tippoint(aes(color=height_max_m, size=10)) +
  scale_color_gradientn(colours=c("blue", "green", "yellow", "red")) +
  geom_tiplab(aes(color=height_max_m, size=10), hjust = -.1) + 
  geom_tiplab2(aes(label=RegenStrategysimplified), align=F, linetype=NA, 
              size=2, offset=4, hjust=0.5) 
p3 + xlim(0, 0.25) + ylim(0, 30)
print(p3)
ggsave("plots/height_max_m.mapped.colours.rectangular.pdf", width = 50, height = 120, limitsize = FALSE)

X = info %>% 
  dplyr::select(AICorrected, height_max_m) %>% 
  as.matrix()
ggplot(X, aes(displ, hwy, colour = class)) + 
  geom_point()

ggtree(beast_tree, aes(color=rate)) +
    scale_color_continuous(low='darkgreen', high='red') +
p4 <- ggtree(tree, aes(color=rate),
             layout='rectangular',
             ladderize = FALSE,
             continuous = 'colour',
             size=2) %<+% info + 
  scale_color_continuous(low='darkgreen', high='red') +
  geom_tippoint(aes(color=Wooddensity.g.cm3..mean, size = 10)) +
  scale_color_gradientn(colours=c("blue", "green", "yellow", "red")) +
  geom_tiplab(aes(color=Wooddensity.g.cm3..mean, ftype="i",fsize=0.5), hjust = -.1)
print(p4)
ggsave("plots/Wooddensity.g.cm3..mean.mapped.colours.pdf", width = 50, height = 120, limitsize = FALSE)

```

Cool, now let's add the multiple associated matrices: Aridity Index, Drought Scale Damage and Maximum tree height
```{r}
library(ggtreeExtra)
library(ggtree)
library(treeio)
library(tidytree)
library(ggstar)
library(ggplot2)
library(ggnewscale)
library(TDbook)

p <- ggtree(tree, layout='circular') %<+% info + 
  geom_tippoint(aes(color=AICorrected)) +
  scale_color_gradientn(colours= rainbow(5))# +
#  geom_tiplab(aes(color=Drought2020Scale.mean, ftype="i",fsize=0.5), hjust = -.1) +
#  geom_tiplab(aes(color=AICorrected, ftype="i",fsize=0.5), hjust = -.1)
print(p)


# p1 <- p + new_scale_fill() %<+% info +
#   geom_fruit(geom=geom_tile,
#              mapping=aes(Drought2020Scale.mean, fill=Drought2020Scale.mean),
#              color = "grey50", offset = 0.04,size = 0.02)+
#   scale_alpha_continuous(range=c(0, 1),
#                          guide=guide_legend(keywidth = 0.3, 
#                                             keyheight = 0.3, order=5))# +
#   geom_fruit(data=dat3, geom=geom_bar,
#              mapping=aes(y=ID, x=HigherAbundance, fill=Sites),
#              pwidth=0.38, 
#              orientation="y", 
#              stat="identity",
#   )

p <- ggtree(tree, layout='circular')

p1 <- p + new_scale_fill() %<+% info +
  geom_fruit(geom=geomtile,
             mapping=aes(fill=Drought2020Scale.mean),
             color = "grey50", offset = 0.04,size = 0.02) +
  
  geom_fruit(geom=geom_tile,
             mapping=aes(color=height_max_m) +
               scale_color_gradientn(colours= c("yellow","pink"))
```

```{r}
circ <- ggtree(tree, layout = "circular")

df <- info %>% 
  select(UniqueID, height_max_m) %>% 
  as.data.frame() %>% 
  glimpse()

df1 <- info %>% 
  select(UniqueID, Drought2020Scale.mean, AICorrected) %>% 
  as.data.frame() %>% 
  glimpse()


p1 <- gheatmap(circ, df, offset=.8, width=.2,
               colnames_angle=95, colnames_offset_y = .25) +
    scale_fill_viridis_d(option="D", name="discrete\nvalue")

test_tree <- p1$data %>% 
  as.phylo()
class(test_tree)

ggtree(test_tree)

library(ggnewscale)
p2 <- p1 + new_scale_fill()
gheatmap(p2, df1, offset=15, width=.3,
         colnames_angle=90, colnames_offset_y = .25) +
    scale_fill_viridis_c(option="A", name="continuous\nvalue")

```

####

librarian::shelf(tidyverse, treeio, ggnewscale, readr, conflicted, ggtree, ape, data.table, phytools, TreeTools,tidytree,  DesiQuintans/desiderata)
conflict_prefer("filter", "dplyr")

tree <- read.tree("plots/final_tree_uniqueID.tre")
info <- read.csv("data/DroughtImpact_SummaryBy_Completetraits.csv")


X = info %>% 
  dplyr::select(AICorrected, height_max_m) %>% 
  as.matrix()
ggplot(X, aes(displ, hwy, colour = class)) + 
  geom_point()

ggtree(tree, aes(color=rate)) +
  scale_color_continuous(low='darkgreen', high='red')

ggtree(tree, layout='rectangular', color="black") %<+% info + 
  geom_tippoint(aes(color=Drought2020Scale.mean, size = 5)) +
  scale_color_gradientn(colours=c("blue", "green", "yellow", "red")) 
ggtree(tree, layout='rectangular', color="black") %<+% info + 
#  geom_tree(aes(color=AICorrected), continuous="color", size=0.2) +
#  scale_color_gradientn(colours=c("blue","cyan","green", "yellow", "orange", "red", "black"))
#  scale_color_gradientn(colours=c("blue", "green", "yellow", "red")) #+
#  scale_color_continuous(low='darkgreen', high='red') +
  geom_tippoint(aes(color=height_max_m, size = 5)) +
  scale_color_gradientn(colours=c("purple","blue","cyan", "green", "yellow", "orange", "red")) #+
#  geom_tiplab(aes(color=Wooddensity.g.cm3..mean, ftype="i",fsize=0.5), hjust = -.1) +

t2 = t1 + new_scale_fill()
  geom_tile(aes(color=AICorrected)) +
  scale_color_gradientn(colours=c("blue", "green", "yellow", "red")) #+

ggtree(p4)
ggsave("plots/Wooddensity.g.cm3..mean.mapped.colours.pdf", width = 50, height = 120, limitsize = FALSE)
